---
title: "Gov 52 Final Project"
author: "Owen Bernstein"
date: "5/4/2021"
output: pdf_document
---

# Introduction
```{r master_script, include = FALSE}

# Loading libraries

library(tidyverse)
library(webshot)
library(gt)
library(readstata13)
library(foreign)
library(data.table)
library(tidyr)
library(stringi)
library(stringr)
library(dplyr)
library(multiwayvcov)
library(lmtest)
library(stargazer)
library(ggplot2)
library(DataCombine)
library(lfe)
library(Matching)
  
output_path <- paste(getwd(),"/figures/",sep="") # stores results to the output file
storewd     <- getwd() # resets working directory when necessary
  
# Sourcing "marginal effects function" script provided by authors

    source("scripts/a.continuous_margeff_function.R")

###################################################################################################
# analysis scripts ################################################################################
###################################################################################################

source("scripts/b.summary_statistics.R")
source("scripts/c.main_analysis.R")
source("scripts/d.legislative_professionalism_analysis.R")
source("scripts/e.extension_analysis.R")
  
```

In their paper “Legislative Term Limits and Polarization,” Michael Olson and Jon
Rogowski explore the causal impact of the adoption of legislative term limits on
representation in state governments.^[Replication data and code for this report
can be accessed on github at https://github.com/OwenBernstein/Gov_52_Final_Project. The orignal report's data and code can be accessed in the Harvard Dataverse at https://doi.org/10.7910/DVN/GDZTK8] ^[Olson, Michael P, & Rogowski, Jon C. (2020). Legislative Term Limits and Polarization. The Journal of Politics, 82(2), 572-586.]
The paper specifically asks how legislative term limits affect representation
and polarization and if this relationship is modified by the professionalism of
the legislature. The authors hypothesize that the implementation of term limits
will increase polarization, and that the increase will be greater in more
professional legislatures. This hypothesis is generated based on two causal
mechanisms:

- Firstly, the authors theorize that term limits reduce a legislator’s need to
understand and respond to their constituency’s demands as they know that they
will not have the possibility of a long career or repeated reelection.

- Secondly, the authors theorize that qualified candidates will be less likely
to run for positions with term limits due to the reduced job security, and as a
result, parties will be more inclined to run candidates with strong ideological
views and party loyalty.

To answer the above questions, the paper uses panel data on roll call voting in
the years between 1993 and 2016 by state as well as indicators for a divided
government, legislative professionalism, and party competitiveness. The primary
model used by the authors is a multiunit difference-in-difference design
described by this linear regression model:

$Y_{it} = \beta_0 + \beta_1\ Term\ limits_{it} + X_{it}\Omega + D_i + T_t + \epsilon_{it}$

In the above regression, Y is the level of polarization in a state's legislature
i in year t, the term limits variable indicates whether state i had term limits
in year t, $\Omega$ is a vector of coefficients for state-level time varied
covariates X~it~ (divided government, legislative professionalism, party
competitiveness), D~i~ is an indicator for state fixed effects, T~t~ is an
indicator for year fixed effects, and $\epsilon$ is a random error term.

The authors concluded that the implementation of term limits led to higher
levels of polarization in state legislatures as evidenced by an increased
ideological gap between the parties in voting patterns. Furthermore, the authors
conclude that this effect is larger for more professional legislatures defined
as those which have longer sessions and higher salaries. The authors also find
that the effects of term limits appeared to be twice as large for Republican
legislators as for Democratic legislators. Finally, these results were robust
across a large number of model types and specifications.

# Replication Analysis

During this project, I was able to replicate all graphs, tables, and models
created in the original paper. This includes both the following replicated
materials as well as the supplementary materials which are not included in this
report. I would like to thank the authors of the original paper, Michael P.
Olson and Jon C. Rogowski, for making their data and code available to
curious students.

## Summary Statistics

In this paper, the authors seek to understand how the implementation of term
limits affects polarization in state legislatures. The main outcome variable is
legislative polarization. This is a measure of polarization devised by Shor and
McCarty (2011) which uses roll call voting to classify each legislator's voting
behavior on an ideological scale that ranges from approximately 0-3, with three being the most polarized.
These individual scores are then aggregated to create a legislative
polarization score for each state-year. The main explanatory variables used
by the authors are term limits, divided government, logged legislative
professionalism, and party competitiveness. Term limits is a binary variable
that takes a value of 1 in state-years in which the legislature had a term
limit. The divided government variable is also a binary variable taking the
value of 1 in divided legislatures. The legislative professionalism variable,
created by Squire (1992) and updated in Squire (2017), is a measure of a
legislature’s professionalism which takes into account salary, staff, and time
in session. Finally, the party competitiveness variable is a measure of the
majority party’s seat share advantage. Below is a summary of each of these
variables, specifically each variable’s mean, median, minimum, maximum, and
standard deviation.

### Recreation of Table A.1: Summary Statistics of Key Variables^[While creating this report I encountered issues with TinyTeX. For this reason, tables in this report are recreated using the "text" option in Stargazer rather than LaTeX.]

```{r taba1, warning=FALSE, echo=F}

sumstat_sg_recreate <- stargazer(pol_dat[,c("l_diffs","house_term_limit","divided_gov",
                                   "legprofscore","abs_diff")],
                                 type = "text",
                        summary.stat=c("mean","median","min","max","sd"),
                        covariate.labels=c("Legislative Polarization","Term Limits","Divided Gov."
                                           ,"ln(Leg. Professionalism)","Party Competitiveness"),
                        title="Summary Statistics of Key Variables",
                        table.layout ="-c-b-",table.placement = "!ht",
                        label="summary_stats",
                        no.space = T)

```

## Main Models

To estimate the effect of term limits on legislative polarization, the authors
use a variety of fixed effect linear models. In each model, the authors use a
combination of the explanatory variables as well as state and year fixed effects
to estimate legislative polarization. By including these fixed effects the
authors seek to account for the exogenous effects caused by state and year. The
effects of state and year are held constant which allows the authors to sort out
the effect of these variables. The main models are described in Table 1.

The baseline model created by the authors (column 1), is a linear regression of
term limits on legislative polarization with fixed effects for both year and
state.

```{r, baseline_model, eval=F}

# Baseline model: l_diffs is legislative polarization, term_limit_temp is term limits

baseline_model <- felm(l_diffs ~ term_limit_temp |state+year|0|state, data=pol_dat)
```

The authors then create a second model in which covariates for divided
government, legislative professionalism, and party competitiveness are added
(column 2).

```{r, extended_model, eval=F}

# Extended model: legprofscore is legislative professionalism, temp_abs_diff is
# party competitiveness

extended_model <- felm(l_diffs ~ term_limit_temp
                       +divided_gov
                       +legprofscore
                       +temp_abs_diff
                   |state+year|0|state
                   ,data=pol_dat)
```

The same baseline and extended models are then applied to exclusively state
legislature Houses (columns 3, 4) and Senates (columns 5,6). The author chose to
apply the model to the two chambers separately as they believed their could be
asymmetric effects across the chambers. This theory was developed based on the
common career pathway of legislators that begin in the lower chambers and then
end in the upper chambers.

### Recreation of Table 1: Term Limits and Polarization

```{r tab1, warning=F, echo=F}

panel_sg_recreate <- stargazer(pooled_nocov,pooled_cov,house_nocov,house_cov,senate_nocov,senate_cov,
                  type = "text",
                  add.lines = list(c("State Fixed Effects","Yes","Yes","Yes","Yes","Yes","Yes"),
                                   c("Year Fixed Effects","Yes","Yes","Yes","Yes","Yes","Yes"),
                                   c("Projected R Squared",round(summary(pooled_nocov)$P.r.squared,3),round(summary(pooled_cov)$P.r.squared,3)
                                     ,round(summary(house_nocov)$P.r.squared,3),round(summary(house_cov)$P.r.squared,3)
                                     ,round(summary(senate_nocov)$P.r.squared,3),round(summary(senate_cov)$P.r.squared,3))),
                  notes.append = FALSE,notes.label = "",
                  notes= "",
                  omit.stat = c("rsq", "f", "ser","adj.rsq", "n"),font.size = "footnotesize",
                  star.char=c("*","**"),star.cutoffs = c(0.10,0.05),digits=3,digits.extra=0,
                  dep.var.caption="Polarization",table.placement = "!ht",
                  dep.var.labels = c("Pooled","House","Senate"),
                  covariate.labels=c("Term Limits","Divided Gov.",
                                     "ln(Leg. Professionalism)","Party Competitiveness"),
                  label="baseline_panel_analysis",
                  table.layout ="-ld-#-tas-n",
                  title="Term Limits and Polarization",
                  out = "figures/panel_sg_recreate.txt",
                  no.space = T)

```

As seen in the table above, the coefficient on term limits is statistically
significant at an alpha level of 0.05 in both the baseline and extended models
for the pooled data and House data, and is not statistically significant for the
Senate data. In the pooled and House models, the effect of term limits is robust
to the additional covariates.

In every model, the effect of term limits is positive, meaning that the
implementation of term limits is correlated with an increase in legislative
polarization on average. The magnitude of this effect is greatest in the lower
chambers (House) and smaller in upper chambers.

Substantively, the authors theorize that the positive relationship between term
limits and legislative polarization has two causes. The first is that
legislators are aware that their term is ending are less inclined to meet
the demands of their constituencies and will instead pursue their own ideological
aims or the aims of their party. The second mechanism is that less-qualified
candidates, and candidates more beholden to their political party, are more
likely to hold office in legislatures with term limits. Both of these causes
would lead to increased polarization in voting patterns.

## Models by Political Party

The authors then sought to identify differential effects of term limits on
legislative polarization for Democrats and Republicans. To do this, the authors
use a second measure of legislative polarization with lower values indicating
more liberal views and higher values indicating more conservative views. The
extended model is recreated for each chamber-party combination. Results are
displayed in Table 2.

### Recreation of Table 2: Term Limits and Asymmetric Polarization: Party Medians

```{r tab2, warning=F, echo=F}

asym_sg_recreate <- stargazer(dem_pooled,dem_house,dem_senate,rep_pooled,rep_house,rep_senate,
                     type = "text",
                     add.lines = list(c("Chamber","Pooled","House","Senate","Pooled","House","Senate"),
                                      c("State Fixed Effects","Yes","Yes","Yes","Yes","Yes","Yes"),
                                      c("Year Fixed Effects","Yes","Yes","Yes","Yes","Yes","Yes"),
                                      c("Projected R Squared",round(summary(dem_pooled)$P.r.squared,3),round(summary(dem_house)$P.r.squared,3)
                                        ,round(summary(dem_senate)$P.r.squared,3),round(summary(rep_pooled)$P.r.squared,3)
                                        ,round(summary(rep_house)$P.r.squared,3),round(summary(rep_senate)$P.r.squared,3))),
                     notes.append = FALSE,notes.label = "",
                     notes="",
                     omit.stat = c("rsq", "f", "ser","adj.rsq"),font.size = "footnotesize",
                     star.char=c("*","**"),star.cutoffs = c(0.10,0.05),digits=3,digits.extra=0,
                     dep.var.labels=c("Democrats","Republicans"),
                     dep.var.caption = c("Party Medians"),table.placement = "!ht",
                     covariate.labels=c("Term Limits","Divided Gov.",
                                        "ln(Leg. Professionalism)","Party Competitiveness"),
                     label="asymmetric_polarization",
                     table.layout ="-ld-#-tas-n",
                     title="Term Limits and Asymmetric Polarization",
                     out = "figures/asym_sg_recreate.txt",
                     no.space = T)

```

As seen above, the coefficient on term limits is negative for all Democrats
indicating that term limits move Democratic legislators to more liberal voting
patterns. On the other hand, the coefficient on term limits is positive for all
Republicans indicating that term limits move Republican legislators to more
conservative voting patterns. Another important insight is that the magnitude of
the coefficient is greater for Republican legislators suggesting that term limits
have a larger effect on Republicans than Democrats. Finally, the coefficient is
only statistically significant at an alpha level of 0.05 for the pooled and
lower chamber models for Republicans. Substantively, the authors argue that these
results show that increased polarization in legislatures is primarily caused by
a shift of Republican legislators to the ideological right.

## Models for Additional Measures 

After determining the effects of term limits on legislative polarization, the
authors sought to determine the effect of term limits on related measures
of a legislature's effectiveness. The specific additional measures the authors
use are party share contributions (the share of campaign contributions coming
from political party committees) and PAC share contributions (the share of
campaign contributions coming from political action committees.)

This analysis also focuses on the effect of legislative professionalism.
Specifically, the below models include an interaction term of legislative
professionalism and term limits. This is included in order to determine if the
effect of term limits differs at different levels of legislative
professionalism. Results are displayed in table 3. 

### Recreation of Table 3: Term Limits, Legislative Professionalism, and Party Influence

```{r tab3, warning=F, echo=F}

mech_prof_sg_recreate <- stargazer(log_legprof,pc_legprof,pac_legprof,
                          type = "text",
                          keep=c("house_term_limit","loglegprof"),
                          add.lines = list(c("Controls","Yes","Yes","Yes"),
                                           c("State Fixed Effects","Yes","Yes","Yes","Yes"),
                                           c("Year Fixed Effects","Yes","Yes","Yes","Yes"),
                                           c("Projected R Squared",round(summary(log_legprof)$P.r.squared,3)
                                             ,paste(as.character(round(summary(pc_legprof)$P.r.squared,3)),"0",sep="")
                                             ,paste(as.character(round(summary(pac_legprof)$P.r.squared,3)),"0",sep=""))),
                          notes.append = FALSE,notes.label = "",
                          notes="",
                          star.char=c("*","**"),star.cutoffs = c(0.10,0.05),
                          omit.stat = c("rsq", "f", "ser","adj.rsq"),font.size = "footnotesize",
                          digits=3,digits.extra=0,
                          dep.var.labels = c("Pooled Polarization","Party Contributions",
                                             "PAC Contributions"),
                          covariate.labels=c("Term Limits","ln(Leg. Professionalism)",
                                             "Term Limits X ln(Professionalism)"),table.placement = "!ht",
                          label="mechanisms_panel",
                          table.layout ="-ld-#-tas-n",
                          title="Term Limits, Legislative Professionalism, and Party Influence",
                          out = "figures/mech_prof_sg_recreate.txt",
                          no.space = T,
                          column.sep.width = "1pt")

```

As seen above, the coefficient on term limits is positive and statistically
significant at an alpha level of 0.05 for the party contributions model. This
suggests that the implementation of term limits is correlated with an increase
in party contributions to campaigns. On the other hand, the share of
contributions that come from PACs decreases with term limits although the
coefficient in this case is not statistically significant at an alpha level of
0.05.

The authors theorize that the relationship between term limits and party
contributions is caused by political parties' desire to fill seats with
ideologically consistent legislators. Specifically, the political parties see
term limits as a way to gain more political influence and consistent voting. On
the other hand, the authors theorize that PACs' main goal is to gain access to
legislators and therefore are less willing to spend money on term-limited
legislators.

The coefficient on the interaction of term limits and legislative
professionalism is positive and statistically significant in both the pooled
polarization model and the party contributions model. This suggests that the
effect of term limits on these two measures is greater in more professional
legislatures.

The authors theorize that this is because legislators in more professional
settings typically have more knowledge and expertise than those in less
professional settings. The implementation of term limits forces these
experienced legislators out of office and replaces them with significantly less
qualified, and typically more ideologically consistent, legislators. This means
that the effect of term limits will have a greater magnitude in professional
legislatures. The interaction between term limits and legislative
professionalism is presented graphically in figure 1.

### Recreation of Figure 1:  Effect of Term Limits over Legislative Professionalism

```{r fig1, warning=F, echo=F, message=F}

logout

```

This graph show that as logged legislative professionalism increases, the
marginal effect of term limits on political polarization also increases. As the
authors theorized, the effect of term limits is greatest in the most professional
legislatures.

## Model Robustness Checks

To confirm the relationship between term limits and political polarization, the
authors run models with additional covariates as a robustness check. The
additional covariates include state level variables such as logged population,
per capita income, unemployment rate, percent foreign born, and Gini
coefficient. Results are presented in Table B.3.

\newpage

```{r, additional_covariates_model, eval=F}

# Additional Covariates Model: 

ec_senate_cov <- felm(s_diffs~term_limit_temp
                      +divided_gov+legprofscore
                      +temp_abs_diff+gov_party
                      +logpop+pc_income
                      +unemploy+pct_fb
                      +gini
                      |state+year|0|state
                      ,data=pol_dat)

```

### Recreation of Table B.3: Term Limits and Polarization: Additional Covariate Control

```{r tabb3, warning=F, echo=F}

extra_controls_sg_recreate <- stargazer(ec_pooled_cov,ec_house_cov,ec_senate_cov,
                                        type = "text",
                               add.lines = list(c("State Fixed Effects","Yes","Yes","Yes","Yes","Yes","Yes"),
                                                c("Year Fixed Effects","Yes","Yes","Yes","Yes","Yes","Yes"),
                                                c("Projected R Squared",round(summary(ec_pooled_cov)$P.r.squared,3)
                                                  ,round(summary(ec_house_cov)$P.r.squared,3)
                                                  ,round(summary(ec_senate_cov)$P.r.squared,3))),
                               notes.append = FALSE,notes.label = "",
                               notes="",
                               omit.stat = c("rsq", "f", "ser","adj.rsq"),font.size = "footnotesize",
                               star.char=c("*","**"),star.cutoffs = c(0.10,0.05),digits=3,digits.extra=0,
                               dep.var.caption="Polarization",table.placement = "!ht",
                               dep.var.labels = c("Pooled","House","Senate"),
                               covariate.labels=c("Term Limits","Divided Gov.","ln(Leg. Professionalism)","Party Competitiveness","Democratic Governor"
                                                  ,"ln(Population)","Per Capita Income","Unemployment Rate"
                                                  ,"Percent Foreign Born","State Gini Coefficient"),
                               label="extra_covariates",
                               table.layout ="-ld-#-t-as-n",
                               title="Term Limits and Polarization: Additional Covariate Control",
                               out = "figures/extra_controls_sg_recreate.txt",
                               no.space = T)

```

As seen above, the coefficient on term limits continues to be positive and
statistically significant at an alpha level of 0.05 for the pooled data and
lower chamber data. This suggests that the effect of term limits is robust to a
variety of additional covariates.

# Extension

To extend the original analysis of the "Legislative Term Limits and
Polarization" paper, I have created two additional models. Each of these models
are fixed effect linear models and the outcome variable continues to be
legislative polarization as used in the previous models. My additional models
differ in their explanatory variables. In the first additional model, I add a
fixed effect for region based on a theory that region effects the performance of
legislators. The second additional model lags the effect of the implementation
of term limits by 4 years to account for the possibility that effects from
implementation may not occur until legislators complete their term in office.
Below, I explain the reasoning for each additional model and display the
results.

## Effect of Region

The first additional model I create adds a fixed effect to the baseline model
created by the authors. The region variable I create takes four values,
Northeast, Midwest, South, and West. There are 10 states in the Northeast, 11 in
the Midwest, 13 in the West, and 15 in the South for a total of 49.^[Nebraska is
excluded as specified in the authors' original report.]

I chose to add a regional fixed effect based on the theory that legislators'
behavior differs based on their geographic region. Kernell (1977) argues that
less legislator turnover in the South has created a more professional
legislature with relatively more expertise. Based upon this assumption, I
hypothesize that implementing term limits in the South would cause a larger
increase in polarization as the skill and qualification gap between the new
legislators and prior legislators would be greater than in other regions. For
this reason, adding a region fixed effect could control for these differences.

In the table below the results of six regression models are displayed. Columns 1
and 2 use pooled data, 3 and 4 use lower chamber data, and 5 and 6 use upper
chamber data. Columns 2, 4, and 6 include a region fixed effect.

### Table 4: Region Fixed Effect

```{r, tab4, warning=F, echo=F}

panel_region_sg_recreate <- stargazer(pooled_nocov,pooled_nocov_region,house_nocov,house_nocov_region,senate_nocov,senate_nocov_region,
                               type = "text",
                               add.lines = list(c("State Fixed Effects","Yes","Yes","Yes","Yes","Yes","Yes"),
                                                c("Year Fixed Effects","Yes","Yes","Yes","Yes","Yes","Yes"),
                                                c("Region Fixed Effects", "No", "Yes", "No", "Yes", "No", "Yes"),
                                                c("Projected R Squared",round(summary(pooled_nocov)$P.r.squared,3),round(summary(pooled_nocov_region)$P.r.squared,3)
                                                  ,round(summary(house_nocov)$P.r.squared,3),round(summary(house_nocov_region)$P.r.squared,3)
                                                  ,round(summary(senate_nocov)$P.r.squared,3),round(summary(senate_nocov_region)$P.r.squared,3))),
                               notes.append = FALSE,notes.label = "",
                               notes= "Note. Linear Regression Coeffcients with standard errors clustered on states in parentheses.",
                               omit.stat = c("rsq", "f", "ser","adj.rsq", "n"),font.size = "footnotesize",
                               star.char=c("*","**"),star.cutoffs = c(0.10,0.05),digits=3,digits.extra=0,
                               dep.var.caption="Polarization",table.placement = "!ht",
                               dep.var.labels = c("Pooled","House","Senate"),
                               covariate.labels=c("Term Limits","Divided Gov.",
                                                  "ln(Leg. Professionalism)","Party Competitiveness"),
                               label="baseline_panel_analysis",
                               table.layout ="-ld-#-tas-n",
                               title="Term Limits and Polarization With Region Fixed Effects",
                               out = "figures/panel_region_sg_recreate.txt",
                               no.space = T)
```

As seen in the table above, the addition of region fixed effects has no effect
on any of the models. The coefficient on term limit remains the same in each
variation of the model regardless of the addition of the region fixed effects.
This suggets that region does not help explain differences in legislative
polarization.

## Lagged Polarization

The second additional model I create lags the term limits variable by 4 years
for each state. The original model attempts to explain legislative polarization
as a function of term limits in the same year. This approach could potentially
ignore the delayed effects of the implementation of term limits. For example,
the implementation of term limits may not immediately effect the behavior of
legislators as they could still run for a single additional term or because the
highly qualified legislators remain in office until the end of their current
term. However, after 4 years, many legislators will have been forced to leave
office or will no longer be allowed to run for additional terms. For this
reason, the effect of term limits may be delayed up to 4 years. Therefore, I
hypothesize that the effect of term limits on legislative polarization will be
greater when using the lagged term limits variable.

The table below shows the results of the additional models. Each column uses
pooled data. Columns 1 and 2 use only term limit as an explanatory variable,
while 3 and 4 use additional covariates. Columns 2 and 4 use the lagged term
limits variable.

### Table 5: Lagged Polarization

```{r, tab5, warning=F, echo=F}

panel_lag_sg_recreate <- stargazer(pooled_nocov,pooled_nocov_lag,pooled_cov,pooled_cov_lag,
                                      type = "text",
                                      add.lines = list(c("State Fixed Effects","Yes","Yes","Yes","Yes"),
                                                       c("Year Fixed Effects","Yes","Yes","Yes","Yes"),
                                                       c("Lag", "No", "Yes", "No", "Yes"),
                                                       c("Projected R Squared",round(summary(pooled_nocov)$P.r.squared,3),round(summary(pooled_nocov_lag)$P.r.squared,3)
                                                         ,round(summary(pooled_cov)$P.r.squared,3),round(summary(pooled_cov_lag)$P.r.squared,3))),
                                      notes.append = FALSE,notes.label = "",
                                      notes= "Note. Linear Regression Coeffcients with standard errors clustered on states in parentheses.",
                                      omit.stat = c("rsq", "f", "ser","adj.rsq", "n"),font.size = "footnotesize",
                                      star.char=c("*","**"),star.cutoffs = c(0.10,0.05),digits=3,digits.extra=0,
                                      dep.var.caption="Polarization",table.placement = "!ht",
                                      dep.var.labels = c("Pooled"),
                                      covariate.labels=c("Term Limits","Divided Gov.",
                                                         "ln(Leg. Professionalism)","Party Competitiveness"),
                                      label="baseline_panel_analysis",
                                      table.layout ="-ld-#-tas-n",
                                      title="Term Limits and Polarization",
                                      out = "figures/panel_lag_sg_recreate.txt",
                                      no.space = T)

```

As seen in the table above, using a lagged term limit variable does affect the
model output. For both the baseline and extended models, the usage of the lagged
term limits variable increased the magnitude of the coefficent on term limits.
This is in line with my hypothesis that the effect of term limits becomes more
pronounced after 4 years. However, the standard error for both the baseline and
extended models also increased. This means that the coefficient on term limits
is no longer significant at an alpha level of 0.05 when using the lagged term
limits variable. This could be because lagging the term limits variable results
in fewer observations of legislative polarization in which term limits have been
implemented.

# Conclusion 

In this project, I was able to replicate all of the relevant graphs, tables, and
models created in the original paper. Specifically, I was able to replicate the
various models used to understand the relationship between the implementation of
term limits and legislative polarization.

This research has shown that the implementation of term limits increases
legislative polarization. This is most likely because legislators are less
inclined to meet the demands of their constituencies and because term limits
lead to the election of less qualified, more ideologically consistent
legislators. Furthermore, this research has shown that these effects are most
prominent in the lower chambers of a state's legislature and also more
pronounced among Republicans. The research also concludes that the relationship
between term limits and legislative polarization is more pronounced in more
professional legislatures. This is most likely because replacing the highly
qualified, expert legislators in professional legislatures with less qualified
legislators has a greater effect than replacing the less qualified legislators in
less professional legislatures. The final conclusion of the paper is that
term limits increase the share of campaign contributions that come from party
committees and decrease the share that come from PACs. This is most likely
because party committees have a greater interest in securing votes while PACs
have an interest in extended access to legislators.

In extending the analysis of this paper, I added a fixed effect for region and
created an additional model with lagged effects of term limits. The regional
fixed effect did not have any effect on the model and therefore suggests that
regions do not help explain differences in legislative polarization. Using the
lagged term limits variable did have an effect on the model. Specifically, the
magnitude of the effect of term limits on legislative polarization was greater
when lagged term limits were used. This suggests that the effect of term limits
is most prominent at least four years after their implementation.

The conclusions of this paper are limited by a variety of factors. The first limiting factor
is that the main outcome variable, legislative polarization, is constant
over the course of a legislators career which means that individual level
changes in polarization can not be included in models. A second potential
limitation could be the existence of reverse causality. In this context, it is
possible that legislatures that are becoming more polarized lead to the
implementation of term limits rather than term limits causing increased
legislative polarization.

Further research on this topic could address the first of these limitations.
Specifically, further research could use a different measure of polarization to
discover how the implementation of term limits affects an individual's voting
behavior rather than the legislative body as a whole. Another avenue for further
research could be voters' responses to term limits. This research could answer
questions such as: do voters respond to legislators' increased polarization?
Future research could also investigate parties', PACs', and activists' responses
to term limits. Aside from changing spending patterns, do these groups behave
differently in the presence of term limits?

Finally, I would once again like to thank the original authors of this paper,
Micael P. Olson and Jon C. Rogowski, for making their data and code available to
other researchers.

# Bibliography

\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\noindent

Olson, Michael P, & Rogowski, Jon C. (2020). Legislative Term Limits and Polarization. The Journal of Politics, 82(2), 572-586.

Peverill Squire. (1992). Legislative Professionalization and Membership Diversity in State Legislatures. Legislative Studies Quarterly, 17(1), 69-79.

Peverill Squire. (2017). A Squire Index Update. State Politics & Policy Quarterly, 17(4), 361-371.

Samuel Kernell. (1977). Toward Understanding 19th Century Congressional Careers: Ambition, Competition, and Rotation. American Journal of Political Science, 21(4), 669-693.

Shor, Boris, & Mccarty, Nolan. (2011). The Ideological Mapping of American Legislatures. The American Political Science Review, 105(3), 530-551.
